\systemnamenospace: Head Orientation Based Selection\\in Physical Spaces

### 2014-06-27 doc/sui14/relatedwork.tex
editing for grammar
fix for pg 2 whitespace

%!TEX root = sui14.tex
\section{Background and Related Work}
Our approach is related to head- and eye-controlled interfaces, area cursors and prior work on hardware devices for pointing in physical spaces.

\vfill
\ben{make sure no section title appears at the end of any column.}
\subsection{Head and Gaze Input}
\changes{Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}. However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.}
%
%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}. 
%
Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.

@@

\subsection{Area Cursors}
\changes{
In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. We argue that head orientation pointing has analogous characteristics (limited pointing performance and accuracy). Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~
\cite{bacim2013design}. In all area and space cursors, the large activation area can lead to multiple targets being selected and disambiguation is needed. This paper describes the trade-offs between several disambiguation approaches.

>>>>

%!TEX root = sui14.tex
\vfill
%\ben{make sure no section title appears at the end of any column.}
\section{Background and Related Work}
Our approach is related to head- and eye-controlled interfaces, area cursors and prior work on hardware devices for pointing in physical spaces.

\subsection{Head and Gaze Input}
\changes{Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assistive input technology for cursor control of desktop applications~\cite{radwin1990method}. However, it is notable that human neck muscles have a lower bandwidth than other muscle groups, e.g., the wrist~\cite{card_morphological_1991}.}
%
%Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse~\cite{Card:1991:MAD:123078.128726}. 
%
Prior work often focused on head orientation for controlling graphical interfaces; in contrast, we apply this modality to selection in physical spaces.

@@

\subsection{Area Cursors}
\changes{
In 2D area cursors for GUIs, the activation area of the cursor is enlarged, which facilitates acquiring smaller targets~\cite{kabbash1995prince}. We argue that head orientation pointing has analogous characteristics (limited pointing performance and accuracy). Area cursors are especially appropriate for individuals with motor control impairments or difficulties~\cite{worden1997making,findlater2010enhanced}. Similar ideas have also been extended into 3D to provide selection with progressive refinement in 3D scenes~
\cite{bacim2013design}. All area and space cursors necessitate disambiguation when there are multiple potential targets and no clear winner. This paper describes the trade-offs between several disambiguation approaches.
